Meta analysis of differential expression results from AMP-AD (Control vs AD, reprocessed RNASeq)

### 2017-02-13 code/Rmd/MetaAnalysis.Rmd
Major modifications

---
title: "Meta analysis of differential expression results from AMP-AD (Control vs AD)"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knit2synapse)
synapseLogin()

knit2synapse::knitToFolderEntity(file = "./MetaAnalysis.Rmd", 
                                 parentId ="syn5569102",
                                 entityName = 'MetaAnalysis of Control vs AD from all three brain banks')
```

@@

# Load required libraries
library(CovariateAnalysis)
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)

library(RColorBrewer)
library(ggplot2)
library(ComplexHeatmap)

library(RankProd)
library(githubr)

@@

library(doParallel)
library(foreach)

clust = makeCluster(14)
registerDoParallel(clust)

@@

### Get covariates/sample metadata from synapse
```{r covariates}
# covariates.id = c('ROSMAP' = 'syn6114444',
#                   'MSBB' = 'syn6131421',
#                   'MAYO' = 'syn6128267')
# SVA Normalised
covariates.id = c('ROSMAP' = 'syn6129368',
                  'MSBB' = 'syn7059296',
                  'MAYO' = 'syn6128630')

covariates = lapply(covariates.id, downloadFile)

covariates$ROSMAP = covariates$ROSMAP %>%
  dplyr::select(Sampleid, msex, cogdx) %>%
  dplyr::rename(ID = Sampleid, Gender = msex, Status = cogdx) %>%
  dplyr::mutate(Gender = factor(Gender, labels = c('1' = 'MALE','2'='FEMALE')),
                Status = factor(Status, labels = c('1' = 'NCI', '2' = 'MCI', '3' = 'MCI.CI', 
                                                   '4' = 'AD', '5' = 'AD.CI', '6' = 'OD')),
                BrainRegion = 'DLPFC') %>%
  dplyr::filter(Status %in% c('NCI', 'AD'))

covariates$MSBB = covariates$MSBB %>%
  dplyr::select(SampleId, SEX, BrainRegion.Diagnosis) %>%
  tidyr::separate(BrainRegion.Diagnosis, c('BrainRegion', 'Dx'), sep = '\\.') %>%
  dplyr::rename(ID = SampleId, Gender = SEX, Status = Dx) %>%
  dplyr::mutate(Gender = factor(Gender, labels = c('M' = 'MALE','F'='FEMALE')),
                Status = factor(Status),
                BrainRegion = factor(BrainRegion, labels = c('BM_10' = 'FP', 'BM_22' = 'STG', 
                                                             'BM_36' = 'PHG', 'BM_44' = 'IFG'))) %>%
  dplyr::filter(Status %in% c('ND', 'SD'))

covariates$MAYO = covariates$MAYO %>%
  dplyr::select(ID, Gender, BrainRegion.Diagnosis) %>%
  tidyr::separate(BrainRegion.Diagnosis, c('BrainRegion', 'Dx'), sep = '\\.') %>%
  dplyr::rename(Status = Dx) %>%
  dplyr::mutate(Gender = factor(Gender, labels = c('M' = 'MALE','F'='FEMALE')),
                Status = factor(Status)) %>%
  dplyr::filter(Status %in% c('Control', 'AD'))

covariates = rbindlist(covariates, use.names = T, fill = T, idcol = 'Study')
```

### Get logCPM counts from synapse
```{r counts}
# logcpm.id = c('ROSMAP' = 'syn6114447' ,
#               'MSBB' = 'syn6131419',
#               'MAYO' = 'syn6128269')
# Get SVA adjusted data
logcpm.id = c('ROSMAP' = 'syn6129370',
              'MSBB' = 'syn7059294',
              'MAYO' = 'syn6128632')

logcpm = lapply(logcpm.id, downloadFile) %>%
  lapply(function(x){
    x = dplyr::select(x, -hgnc_symbol)
    }) %>%
  join_all(type = 'full') %>%
  dplyr::select(Gene.ID, ensembl_gene_id, one_of(covariates$ID)) 
logcpm[is.na(logcpm)] = dplyr::select(logcpm, -Gene.ID, -ensembl_gene_id) %>%
  min
```

### Get residual expression from synapse
```{r counts1}
# Get SVA adjusted data
resid.expr.id = c('ROSMAP' = 'syn6129372',
                  'MSBB' = 'syn7059289',
                  'MAYO' = 'syn6128634')

@@

### Download differential expression data
```{r download.data}
# diffexp.id = c('ROSMAP' = 'syn6114453',
#                'MSBB' = 'syn6131423',
#                'MAYO' = 'syn6128273')
# Get SVA adjusted differential expression
diffexp.id = c('ROSMAP' = 'syn6129374',
               'MSBB' = 'syn7059298',
               'MAYO' = 'syn6128636')

diffexp = lapply(diffexp.id, downloadFile)

diffexp$ROSMAP = diffexp$ROSMAP %>%
  filter(Comparison == 'cogdx1-cogdx4') %>%
  dplyr::select(ensembl_gene_id, logFC, P.Value, adj.P.Val) %>%
  # dplyr::select(Gene.ID, logFC, P.Value, adj.P.Val) %>%
  dplyr::mutate(Tissue = 'DLPFC', Contrast = 'NCI-AD')

diffexp$MSBB = diffexp$MSBB %>%
  dplyr::filter(Name == 'ND-SD') %>%
  dplyr::select(ensembl_gene_id, logFC, P.Value, adj.P.Val, BrainRegion, Name) %>%
  dplyr::mutate(BrainRegion = factor(BrainRegion, labels = c('BM_10' = 'FP', 'BM_22' = 'STG', 
                                                             'BM_36' = 'PHG', 'BM_44' = 'IFG'))) %>%
  dplyr::rename(Tissue = BrainRegion, Contrast = Name)

diffexp$MAYO = diffexp$MAYO %>%
  dplyr::filter(Name == 'Control-AD') %>%
  tidyr::separate(Gene.ID, c('ensembl_gene_id', 'position'), sep = '\\.') %>%
  dplyr::select(ensembl_gene_id, logFC, P.Value, adj.P.Val, BrainRegion, Name) %>%
  dplyr::rename(Tissue = BrainRegion, Contrast = Name)

diffexp = rbindlist(diffexp, use.names = T, fill = T, idcol = 'Study')

diffexp$DE = 'NONE'
diffexp$DE[diffexp$adj.P.Val <= 0.05 & diffexp$logFC >= log2(1.2)] = 'UP'
diffexp$DE[diffexp$adj.P.Val <= 0.05 & diffexp$logFC <= -log2(1.2)] = 'DOWN'

pl = list()
for (study in c('ROSMAP', 'MSBB', 'MAYO')){
  p = filter(diffexp, Study == study) %>%
    ggplot(aes(x = logFC, y = -log10(adj.P.Val), color = DE)) 
  p = p + scale_colour_manual(values = c("green","grey", "red"))
  p = p + geom_point() + theme(legend.position = 'None')
  p = p + facet_grid(Study~.+Tissue)
  pl[[study]] = p
}
multiplot(plotlist = pl, cols = 1)
```

### Get hgnc_symbol mapping from biomaRt
```{r biomart}
# Define biomart object
mart <- useMart('ENSEMBL_MART_ENSEMBL', dataset = "hsapiens_gene_ensembl", host = 'dec2015.archive.ensembl.org')

# Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"),
                       filters = "ensembl_gene_id", 
                       values = unique(diffexp$ensembl_gene_id),
                       mart = mart) %>%
  data.table()

# Add hgnc_symbols to the logcpm matrix
logcpm = logcpm %>%
  left_join(Ensemble2HGNC)

# Add hgnc_symbols to the resid.mat matrix
resid.expr = resid.expr %>%
  left_join(Ensemble2HGNC)

# Add hgnc_symbols to the diffexp matrix
diffexp = diffexp %>%
  left_join(Ensemble2HGNC)

diffexp$Direction = 'UP'
diffexp$Direction[diffexp$logFC < 0] = 'DOWN'

writeLines('Differentially expressed at an fdr of 0.05')
diffexp %>%
  filter(adj.P.Val <= 0.05) %>%
  group_by(Tissue, Direction) %>%
  summarise(n = length(unique(ensembl_gene_id))) %>%
  spread(Direction, n)

writeLines('Differentially expressed at an fdr of 0.05 and fold change of 1.2')
diffexp %>%
  filter(adj.P.Val <= 0.05, (logFC >= log2(1.2) | logFC <= -log2(1.2))) %>%
  group_by(Tissue, Direction) %>%
  summarise(n = length(unique(ensembl_gene_id))) %>%
  spread(Direction, n)
```

### Pathway enrichment
```{r pathway.enrichment}
# Differentially expressed in ROSMAP
## Up regulated
tmp = filter(diffexp, adj.P.Val <= 0.05, logFC >0, Tissue == 'DLPFC') %>%
  dplyr::left_join(Ensemble2HGNC %>% as.data.table()) %>%
  filter(hgnc_symbol != '')
### Enrichr results are here: http://amp.pharm.mssm.edu/Enrichr/enrich?dataset=6x3h

## Down regulated
tmp = filter(diffexp, adj.P.Val <= 0.05, logFC<0, Tissue == 'DLPFC') %>%
  dplyr::left_join(Ensemble2HGNC %>% as.data.table()) %>%
  filter(hgnc_symbol != '')
### Enrichr results are here: http://amp.pharm.mssm.edu/Enrichr/enrich?dataset=6x4e

# Differentially expressed in MSSM FP
## Up regulated
tmp = filter(diffexp, adj.P.Val <= 0.05, logFC >0, Tissue == 'FP') %>%
  dplyr::left_join(Ensemble2HGNC %>% as.data.table()) %>%
  filter(hgnc_symbol != '')
### Enrichr results are here: http://amp.pharm.mssm.edu/Enrichr/enrich?dataset=6xcm

## Down regulated
tmp = filter(diffexp, adj.P.Val <= 0.05, logFC < 0, Tissue == 'FP') %>%
  dplyr::left_join(Ensemble2HGNC %>% as.data.table()) %>%
  filter(hgnc_symbol != '')

# Differentially expressed in MSSM STG
## Up regulated
tmp = filter(diffexp, adj.P.Val <= 0.05, logFC > 0, Tissue == 'STG') %>%
  dplyr::left_join(Ensemble2HGNC %>% as.data.table()) %>%
  filter(hgnc_symbol != '')

## Down regulated
tmp = filter(diffexp, adj.P.Val <= 0.05, logFC < 0, Tissue == 'STG') %>%
  dplyr::left_join(Ensemble2HGNC %>% as.data.table()) %>%
  filter(hgnc_symbol != '')

# Differentially expressed in MSSM PHG
## Up regulated
tmp = filter(diffexp, adj.P.Val <= 0.05, logFC > 0, Tissue == 'PHG') %>%
  dplyr::left_join(Ensemble2HGNC %>% as.data.table()) %>%
  filter(hgnc_symbol != '')
### Enrichr results are here: http://amp.pharm.mssm.edu/Enrichr/enrich?dataset=6xco

## Down regulated
tmp = filter(diffexp, adj.P.Val <= 0.05, logFC < 0, Tissue == 'PHG') %>%
  dplyr::left_join(Ensemble2HGNC %>% as.data.table()) %>%
  filter(hgnc_symbol != '')
### Enrichr results are here: http://amp.pharm.mssm.edu/Enrichr/enrich?dataset=6xge

# Differentially expressed in MAYO CER
## Up regulated
tmp = filter(diffexp, adj.P.Val <= 0.05, logFC > 0, Tissue == 'CER') %>%
  dplyr::left_join(Ensemble2HGNC %>% as.data.table()) %>%
  filter(hgnc_symbol != '')
### Enrichr results are here:

## Down regulated
tmp = filter(diffexp, adj.P.Val <= 0.05, logFC < 0, Tissue == 'PHG') %>%
  dplyr::left_join(Ensemble2HGNC %>% as.data.table()) %>%
  filter(hgnc_symbol != '')
### Enrichr results are here: http://amp.pharm.mssm.edu/Enrichr/enrich?dataset=6xge
```

### Meta Analysis
```{r metaanalysis}
P.Value = diffexp %>% 
  # filter(Tissue %in% c('DLPFC', 'PHG', 'TCX')) %>%
  filter(ensembl_gene_id != '') %>%
  dplyr::select(ensembl_gene_id, P.Value, Tissue, Contrast) %>%
  tidyr::unite(Tissue.Contrast, Tissue, Contrast, sep = '_') %>%
  tidyr::spread(Tissue.Contrast, P.Value)
P.Value[is.na(P.Value)] = 1

min.pval = dplyr::select(P.Value, -ensembl_gene_id) %>%
  data.matrix %>% rowMin %>%
  p.adjust(method = 'BH') %>%
  data.frame(ensembl_gene_id = P.Value$ensembl_gene_id,
             min.pval = .)

max.pval = dplyr::select(P.Value, -ensembl_gene_id) %>%
  data.matrix %>% rowMax %>%
  p.adjust(method = 'BH') %>%
  data.frame(ensembl_gene_id = P.Value$ensembl_gene_id,
             max.pval = .)

fb.pval = dplyr::select(P.Value, -ensembl_gene_id) %>%
  data.matrix %>% fishersPvalue %>%
  p.adjust(method = 'BH') %>%
  data.frame(ensembl_gene_id = P.Value$ensembl_gene_id,
             fb.pval = .)

# Rank prod based analysis
expr = resid.expr[!duplicated(resid.expr$ensembl_gene_id),-c(1:2)]
rownames(expr) = resid.expr$ensembl_gene_id[!duplicated(resid.expr$ensembl_gene_id)]

>>>>

---
title: "Meta analysis of differential expression results from AMP-AD (Control vs AD, reprocessed RNASeq)"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knit2synapse)
synapseLogin()

knit2synapse::knitToFolderEntity(file = "./MetaAnalysis.Rmd", 
                                 parentId ="syn5569102",
                                 entityName = 'MetaAnalysis of Control vs AD from all three brain banks (Reprocessed RNASeq)')
```

@@

# Load required libraries
library(CovariateAnalysis) # install from th1vairam repo using devtools::install_github('th1vairam/CovariateAnalysis@devtools')
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)

library(RColorBrewer)
library(ggplot2)
library(ComplexHeatmap)
library(biomaRt)
library(Matrix)
library(org.Hs.eg.db)

library(limma)
library(edgeR)
library(RankProd)
library(githubr)

@@

library(doParallel)
library(foreach)
library(doParallel)

clust = makeCluster(detectCores() - 2)
registerDoParallel(clust)

@@

### Get covariates/sample metadata from synapse
```{r covariates}
covariates.ids = c(ROSMAP = 'syn8018344',
                   MSBB = 'syn8073656',
                   MAYO = 'syn8028561')
covariates = lapply(covariates.ids, function(id){
  x = downloadFile(id)
  colnames(x)[1] = 'SampleID'
  return(x)
})

covariates$ROSMAP = dplyr::filter(covariates$ROSMAP, cogdx %in% c(1,4)) %>%
  dplyr::mutate(Dx = factor(cogdx, levels = c(1,4), labels = c('1' = 'Control', '4' = 'AD')),
                Study = 'ROSMAP',
                BrainRegion = 'DLPFC') %>%
  dplyr::select(SampleID, Study, BrainRegion, Dx)

covariates$MSBB = dplyr::filter(covariates$MSBB, BrodmannArea.Dx %in% c('BM10.SD', 'BM22.SD', 'BM36.SD', 'BM44.SD', 'BM36.ND', 'BM10.ND', 'BM44.ND', 'BM22.ND')) %>%
  dplyr::select(SampleID, BrodmannArea.Dx) %>%
  tidyr::separate(BrodmannArea.Dx, c('BrainRegion', 'Dx')) %>%
  dplyr::mutate(Study = 'MSSM',
                Dx = factor(Dx, levels = c('ND','SD'), labels = c('ND' = 'Control', 'SD' = 'AD')),
                BrainRegion = factor(BrainRegion, labels = c('FP', 'STG', 'PHG', 'IFG'))) %>%
  dplyr::select(SampleID, Study, BrainRegion, Dx)

covariates$MAYO = dplyr::filter(covariates$MAYO, BrainRegion.Diagnosis %in% c('TCX.AD', 'TCX.Control', 'CER.AD', 'CER.Control')) %>%
  dplyr::select(SampleID, BrainRegion.Diagnosis) %>%
  tidyr::separate(BrainRegion.Diagnosis, c('BrainRegion', 'Dx')) %>%
  dplyr::mutate(Study = 'MAYO',
                Dx = factor(Dx, levels = c('Control','AD')),
                BrainRegion = factor(BrainRegion, levels = c('CER', 'TCX'))) %>%
  dplyr::select(SampleID, Study, BrainRegion, Dx)

covariates = covariates %>%
  rbindlist(use.names = T, fill = T)
```

### Get filtered gene.id and sample.id from synapse
```{r counts}
logcpm.ids = c('ROSMAP' = 'syn8018352',
               'MSBB' = 'syn8073666',
               'MAYO' = 'syn8028565')

logcpm = lapply(logcpm.ids, downloadFile) %>%
  lapply(function(x){
    x = dplyr::select(x, -hgnc_symbol, -ensembl_gene_id)
  }) %>%
  join_all(type = 'full') %>%
  dplyr::select(gene.id, one_of(covariates$SampleID)) 
logcpm[is.na(logcpm)] = min(logcpm[,-(1)], na.rm = T)
```

### Get raw counts from synapse
```{r raw.counts, eval = F}
raw.counts.ids = c('MAYO_CBE' = 'syn7513807',
                   'MSBB' = 'syn7994853',
                   'MAYO_TCX' = 'syn7987765',
                   'ROSMAP' = 'syn7750270')

lcpm = lapply(raw.counts.ids, function(id, lcpm){
  expr = read.table(synGet(id)@filePath, sep = '\t', check.names = F) 
  lib.size = calcNormFactors(expr, method = 'TMM')
  expr = voom(expr, lib.size = lib.size, design = NULL, plot = F, normalize.method = 'quantile')$E
  expr = expr[row.names(expr) %in% lcpm$gene.id, ] %>%
    rownameToFirstColumn('gene.id')
}, logcpm) %>%
  join_all(type = 'full')
```

### Get residual expression from synapse
```{r resid.expr}
# Get SVA adjusted data
resid.expr.id = c('ROSMAP' = 'syn6129372',
                  'MSBB' = 'syn7059289',
                  'MAYO' = 'syn6128634')

@@

### Get hgnc_symbol mapping from biomaRt
```{r biomart}
# Define biomart object
mart <- useMart(biomart = "ENSEMBL_MART_ENSEMBL", dataset = "hsapiens_gene_ensembl", host = "dec2016.archive.ensembl.org")

# Get background gene ids
backgroundGenes = logcpm[,'gene.id', drop = F] %>% 
  dplyr::mutate(ID = gene.id) %>%
  tidyr::separate(ID, c('ensembl_gene_id','position'), sep = '\\.')

# Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"),
                       filters = "ensembl_gene_id", values = backgroundGenes$ensembl_gene_id,
                       mart = mart)

Entrez2Ensembl <- as.data.frame(org.Hs.egENSEMBL2EG)
Entrez2HGNC <- as.data.frame(org.Hs.egSYMBOL2EG)
HGNC2Ensembl <- left_join(Entrez2HGNC, Entrez2Ensembl)
HGNC2Ensembl = HGNC2Ensembl[grep('^LOC',HGNC2Ensembl$symbol),c('ensembl_id', 'symbol')] %>%
  filter(!is.na(ensembl_id), !is.na(symbol)) %>%
  dplyr::rename(ensembl_gene_id = ensembl_id, hgnc_symbol = symbol)

Ensemble2HGNC = left_join(backgroundGenes,
                          rbindlist(list(Ensemble2HGNC, HGNC2Ensembl), use.names = T, fill = T) %>%
                            unique) %>%
  unique
```

### Get differential expression results from synapse
```{r diffexp, fig.height=16, fig.width=10}
diff.exp.ids = c(ROSMAP = 'syn8018358',
                 MSBB = 'syn8073678',
                 MAYO = 'syn8028572')
diff.exp = lapply(diff.exp.ids, downloadFile)

# Filter case-control differences alone
diff.exp$ROSMAP = filter(diff.exp$ROSMAP, Comparison == 'cogdx1-cogdx4')
diff.exp$MSBB = filter(diff.exp$MSBB, Comparison == 'ND-SD')
diff.exp$MAYO = filter(diff.exp$MAYO, Comparison == 'Control-AD')

writeLines('Total number of genes differentially expressed between Control-AD at an FDR of 0.05 and foldchange of 1.2 are')
lapply(diff.exp, function(x){
  x = filter(x, adj.P.Val <= 0.05, abs(logFC) >= log2(1.2)) 
  x$Direction = 'NONE'
  x$Direction[x$logFC > 0] = 'UP'
  x$Direction[x$logFC < 0] = 'DOWN'
  x = x %>% group_by(Study, Region, Direction) %>%
    summarise(count = length(unique(gene.id)))
}) %>%
  rbindlist() %>%
  spread(Direction, count) %>%
  kable

tmp = diff.exp %>% 
  rbindlist(use.names = T, fill = T) 
tmp$Direction = 'NONE'
tmp$Direction[tmp$logFC > log2(1.2) & tmp$adj.P.Val <= 0.05] = 'UP'
tmp$Direction[tmp$logFC < -log2(1.2) & tmp$adj.P.Val <= 0.05] = 'DOWN'

pl = list()
for (study in c('ROSMAP','MSBB','MAYO')){
  p = ggplot(tmp %>% filter(Study == study) %>% droplevels, aes(x = logFC, y = -log10(adj.P.Val), color = Direction)) + geom_point()
  p = p + scale_color_manual(values = c('green','grey','red'))
  p = p + facet_grid(Study~.+Region) + theme_bw() + theme(legend.position = 'top')
  pl[[study]] = p
}
multiplot(plotlist = pl, cols = 1)

# Find genes that are differentially expressed in all brain regions
diff.exp.gl = dlply(tmp, .(Region), .fun = function(x){
  dplyr::filter(x, abs(logFC) >= log2(1.2) & adj.P.Val <= 0.05) %>%
    dplyr::select(gene.id) %>%
    unique %>% unlist
})

tmp1 = filter(tmp, gene.id %in% Reduce(intersect, diff.exp.gl)) %>%
  dplyr::select(-ensembl_gene_id, -hgnc_symbol) %>%
  left_join(Ensemble2HGNC %>% dplyr::select(-position))
tmp1$hgnc_symbol[tmp1$ensembl_gene_id == 'ENSG00000275830'] = 'LOC101927646'

tmp.eff.sz = group_by(tmp1, gene.id) %>% summarise(es = sum(logFC, na.rm = T)) %>% top_n(6, abs(es))

writeLines('Top six differentially expressed genes are')
pl = list()
p = ggplot(tmp1 %>% filter(gene.id %in% tmp.eff.sz$gene.id[tmp.eff.sz$es>0]), aes(x = Region, y = logFC, fill = Study)) + geom_bar(stat = 'identity')
p = p + facet_grid(hgnc_symbol~.) + geom_errorbar(aes(ymin=CI.L, ymax=CI.R), width = 0.2)
p = p + theme(legend.position = 'top')
pl[[1]] = p

p = ggplot(tmp1 %>% filter(gene.id %in% tmp.eff.sz$gene.id[tmp.eff.sz$es<0]), aes(x = Region, y = logFC, fill = Study)) + geom_bar(stat = 'identity')
p = p + facet_grid(hgnc_symbol~.) + geom_errorbar(aes(ymin=CI.L, ymax=CI.R), width = 0.2)
p = p + theme(legend.position = 'top')
pl[[2]] = p

multiplot(plotlist = pl, cols = 2)

# Genes differentially expressed in all brain regions
tmp1 = dlply(tmp, .(Region), .fun = function(x){
  dplyr::filter(x, abs(logFC) >= log2(1) & adj.P.Val <= 0.05) %>%
    dplyr::select(gene.id) %>%
    unique %>% unlist
}) %>%
  Reduce(intersect, .)
writeLines('Genes that are differentially expressed between Control-AD at FDR 0.05 in all brain regions')
kable(data.frame(ngenes = length(tmp1),
                 hgnc_symbol = paste(unique(Ensemble2HGNC$hgnc_symbol[Ensemble2HGNC$gene.id %in% tmp1]), collapse = ', ')))

tmp1 = dlply(tmp, .(Region), .fun = function(x){
  dplyr::filter(x, abs(logFC) >= log2(1.2) & adj.P.Val <= 0.05) %>%
    dplyr::select(gene.id) %>%
    unique %>% unlist
}) %>%
  Reduce(intersect, .)
writeLines('Genes that are differentially expressed between Control-AD at FDR 0.05 and foldchange of 1.2 in all brain regions')
kable(data.frame(ngenes = length(tmp1),
                 hgnc_symbol = paste(unique(Ensemble2HGNC$hgnc_symbol[Ensemble2HGNC$gene.id %in% tmp1]), collapse = ', ')))
```

### Meta Analysis
```{r metaanalysis.min.pval}
# Get raw p.values
P.Value = diff.exp %>% 
  rbindlist(use.names = T, fill = T) %>%
  dplyr::select(gene.id, P.Value, Region) %>%
  unique() %>%
  tidyr::spread(Region, P.Value)
P.Value[is.na(P.Value)] = 1
rownames(P.Value) = P.Value$gene.id
P.Value$gene.id = NULL
P.Value = data.matrix(P.Value)

# Perform minimum p-value based meta-analysis
min.pval = P.Value %>% 
  apply(1, min, na.rm = T) %>%
  p.adjust(method = 'bonferroni') %>%
  rownameToFirstColumn('gene.id') %>%
  dplyr::rename(min.pval = DF)

# Perform fisher p-value based meta-analysis
fb.pval = P.Value %>%
  fishersPvalue %>%
  p.adjust(method = 'bonferroni') %>%
  rownameToFirstColumn('gene.id') %>%
  dplyr::rename(fb.pval = DF)
```

```{r }
# Rank prod based analysis
expr = resid.expr[!duplicated(resid.expr$ensembl_gene_id),-c(1:2)]
rownames(expr) = resid.expr$ensembl_gene_id[!duplicated(resid.expr$ensembl_gene_id)]
